---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
```{r}
library(tidyverse)
library(lubridate)
```
## Loading and preprocessing the data
```{r echo=TRUE}
data <- tbl_df(read.csv("activity.csv"))
data <- data %>% mutate(date=as_date(date))
```

## What is mean total number of steps taken per day?
```{r echo=TRUE}
tmp <- data %>% 
  group_by(date) %>% 
  summarize(sum(steps, na.rm=TRUE))
names(tmp) <- c("date","total_steps")
hist(tmp$total_steps, main="Daily Total Steps", xlab="Total Steps", breaks=5)
mean_steps <- mean(tmp$total_steps, na.rm = TRUE)
median_steps <- median(tmp$total_steps, na.rm = TRUE)
```
The mean number of steps is `r format(mean_steps, scientific=FALSE)` and the median number of steps is `r format(median_steps, scientific=FALSE)`.

## What is the average daily activity pattern?
Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r echo=TRUE}
data <- data %>% 
  ungroup() %>% 
  group_by(interval)
tmp <- summarize(data, mean(steps, na.rm = TRUE))
names(tmp) <- c("interval", "avgsteps")
ggplot(tmp, aes(interval, avgsteps)) + 
  geom_line(size=0.8) + 
  xlab("Interval") + 
  ylab("Average Steps")
highest_interval <- top_n(tmp, 1, avgsteps)
```
The interval `r highest_interval$interval` has the highest average steps taken at `r highest_interval$avgsteps`.


## Imputing missing values
Note that there are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data.

1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r echo=TRUE}
total_na <- sum(is.na(data))
```
Total missing values are `r total_na`

2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

Create a new dataset that is equal to the original dataset but with the missing data filled in.

**Strategy:**
We shall be imputing by taking the mean of steps for each interval and replacing the NAs with their respective values.

```{r echo=TRUE}
impute.mean <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))
data_imputed <- data %>% 
  mutate(steps = as.integer(impute.mean(steps)))
```
3. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

```{r echo=TRUE}
data_imputed <- ungroup(data_imputed) %>%
  group_by(date)
data_imputed <- data_imputed %>%
  mutate(daily_steps=sum(steps))
hist(data_imputed$daily_steps, breaks = 5, main="Imputed Total Daily Steps", xlab="Total Daily Steps")
imputed_steps_mean <- mean(data_imputed$daily_steps)
imputed_steps_median <- median(data_imputed$daily_steps)
```
For the imputed dataset the mean is `r format(imputed_steps_mean, scientific=FALSE)` and the median is `r format(imputed_steps_median, scientific=FALSE)`. While for the raw dataset the mean was `r format(mean_steps, scientific=FALSE)` and the median was `r format(median_steps, scientific=FALSE)`.

## Are there differences in activity patterns between weekdays and weekends?
```{r echo=TRUE}
data_imputed <- data_imputed %>% 
  mutate(day_type = ifelse(wday(date, label=TRUE) %in% c("Sat","Sun"), "weekend", "weekday"))

data_imputed <- data_imputed %>% 
  ungroup() %>% 
  mutate(day_type=as_factor(day_type))

tmp <- data_imputed %>% 
  group_by(day_type, interval) %>% 
  summarize(msteps= mean(steps))

ggplot(data=tmp) + 
  aes(x=interval, y=msteps) + 
  geom_line() + 
  facet_grid(rows=vars(day_type)) +
  ylab("Number of Steps") + 
  xlab("Interval")

```


